---
title: "**S-LDSC analysis to estimate the enrichment of GWAS variants in human cell m6A peaks**"
author: "Christie Du"
date: "2020-08-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

I ran S-LDSC and on m6A annotations of 3 cell types: MOLM13, MONO-MAC-6, and NOMO-1. I did this for 11 GWAS: Asthma, BMI, CAD, HDL, Height, Intelligence, IBD, LDL, RA, Schizophrenia, and Type II Diabetes. M6A annotations used are from here: https://repicmod.uchicago.edu/repic/download.php.

## **S-LDSC**

For S-LDSC, I used baselineLD model v1.1: 

1. BaselineLD v1.1 located here:
`RCC:/project2/xinhe/kevinluo/ldsc/LDSCORE/1000G_Phase3_baselineLD_v1.1_ldscores`

### GWAS and m6A annotations:
* GWAS summary statistics in the sumstats format (excluding Intelligence) are located at `RCC:/project2/xinhe/kevinluo/GWAS/GWAS_summary_stats/GWAS_collection/ldsc_format` 
* Intelligence GWAS in the sumstats format is located at `RCC:/project2/xinhe/kevinluo/GWAS/GWAS_summary_stats/GWAS_from_Min/ldsc_format` 
* The original m6A annotations are located at `RCC:/project2/xinhe/cdu/ldsc/annot/m6A_immune_cells/RAW` 

Because the GWAS summary statistics were based on the hg19 map, the m6A annotations were converted from hg38 to hg19 using `liftOver`. The final BED files are located here: `RCC:/project2/xinhe/cdu/ldsc/annot/m6A_immune_cells/hg19`.

```{r annotations and traits}
annot_list <- c("MOLM13", "MONO-MAC-6", "NOMO-1")
trait_name_list <- c("Asthma", "BMI","CAD","HDL","Height","IBD","Intelligence","LDL","RA","Schizophrenia","T2D")
```


```{r plot functions}
library(ggplot2)
library(foreach)
library(doParallel)

ggplot_enrichment <- function(result_sLDSC, xlim = NULL, title = "Enrichment"){
  
  Enrichment <- result_sLDSC$Enrichment
  Enrichment_CI_L <- result_sLDSC$Enrichment - 1.96*result_sLDSC$Enrichment_std_error
  Enrichment_CI_H <- result_sLDSC$Enrichment + 1.96*result_sLDSC$Enrichment_std_error

  ## truncate at 1
  Enrichment[Enrichment < 1] <- 1
  Enrichment_CI_L[Enrichment_CI_L < 1] <- 1
  Enrichment_CI_H[Enrichment_CI_H < 1] <- 1
  
  p <- ggplot(result_sLDSC, aes(x = Enrichment, y = Category, colour = Color))+
    geom_point()+
    xlab("Enrichment")+
    ggtitle(title)+
    geom_errorbarh(aes(xmin = Enrichment - 1.96*Enrichment_std_error, 
                       xmax = Enrichment + 1.96*Enrichment_std_error, height = 0.1))+ 
    facet_wrap(Disease~.,ncol = 3)+
    theme_bw()  + 
    geom_vline(xintercept = 1,linetype="dotted", colour = "red")+
    theme(axis.ticks = element_blank(),  
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          axis.text = element_text(face="bold",size = 10, colour = "black"),
          axis.title = element_text(face="bold",size = 10),
          strip.text = element_text(face="bold",size = 10), 
          panel.spacing.x = unit(0.6,units = "cm"), 
          axis.title.y = element_blank(), 
          legend.position = "none", 
          plot.title = element_text(hjust = 0.5))
  if(!is.null(xlim)){
    p <- p + coord_cartesian(xlim = xlim)
  }
  print(p)
}

ggplot_log2_enrichment <- function(result_sLDSC, xlim = NULL, title = "Enrichment"){
  
  result_sLDSC$Enrichment_CI_L <- result_sLDSC$Enrichment - 1.96*result_sLDSC$Enrichment_std_error
  result_sLDSC$Enrichment_CI_H <- result_sLDSC$Enrichment + 1.96*result_sLDSC$Enrichment_std_error

  ## truncate at 1
  result_sLDSC$Enrichment[result_sLDSC$Enrichment < 1] <- 1
  result_sLDSC$Enrichment_CI_L[result_sLDSC$Enrichment_CI_L < 1] <- 1
  result_sLDSC$Enrichment_CI_H[result_sLDSC$Enrichment_CI_H < 1] <- 1

  p <- ggplot(result_sLDSC, aes(x = log2(Enrichment), y = Category, colour = Color))+
    geom_point()+
    xlab("log2(Enrichment)")+
    ggtitle(title)+
    geom_errorbarh(aes(xmin = log2(Enrichment_CI_L), 
                       xmax = log2(Enrichment_CI_H), height = 0.1))+ 
    facet_wrap(Disease~.,ncol = 3)+
    theme_bw()  + 
    geom_vline(xintercept = 0,linetype="dotted", colour = "red")+
    theme(axis.ticks = element_blank(),  
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          axis.text = element_text(face="bold",size = 10, colour = "black"),
          axis.title = element_text(face="bold",size = 10),
          strip.text = element_text(face="bold",size = 10), 
          panel.spacing.x = unit(0.6,units = "cm"), 
          axis.title.y = element_blank(), 
          legend.position = "none", 
          plot.title = element_text(hjust = 0.5))
  if(!is.null(xlim)){
    p <- p + coord_cartesian(xlim = xlim)
  }
  print(p)
}

## combine S-LDSC enrichment results across traits
combine_sldsc_traits <- function(trait_name_list, dir_results, baseline){
  registerDoParallel(cores = 6)
  
  result_sLDSC <- foreach(trait = trait_name_list, .combine = rbind)%dopar%{
    sldsc_results <- read.table(paste0(dir_results,"/", trait, "/", baseline, "/", trait,"_", "MOLM13", "_hg19_", baseline, ".results"), header = T, stringsAsFactors = F)
    sldsc.MOLM13 <- sldsc_results[sldsc_results$Category == "L2_1",]
    sldsc.MOLM13$Category <- "MOLM13"
    
    sldsc_results <- read.table(paste0(dir_results,"/", trait, "/", baseline, "/", trait,"_", "MONO-MAC-6", "_hg19_", baseline, ".results"), header = T, stringsAsFactors = F)
    sldsc.MONOMAC6 <- sldsc_results[sldsc_results$Category == "L2_1",]
    sldsc.MONOMAC6$Category <- "MONO-MAC-6"
    
    sldsc_results <- read.table(paste0(dir_results,"/", trait, "/", baseline, "/", trait,"_", "NOMO-1", "_hg19_", baseline, ".results"), header = T, stringsAsFactors = F)
    sldsc.NOMO1 <- sldsc_results[sldsc_results$Category == "L2_1",]
    sldsc.NOMO1$Category <- "NOMO-1"
    
    sldsc.combined <- rbind(sldsc.MOLM13, sldsc.MONOMAC6, sldsc.NOMO1)
    sldsc.combined <- cbind(Disease = trait, sldsc.combined)
    sldsc.combined
  }
  return(result_sLDSC)
}
```


```{r v1.1 data import}
baseline <- "baselineLDv1.1"

dir_results <- "/project2/xinhe/cdu/ldsc/results/immune_cells"
result_sLDSC <- combine_sldsc_traits(trait_name_list, dir_results, baseline)

result_sLDSC$Category <- factor(result_sLDSC$Category, levels = rev(annot_list) )
result_sLDSC$Color <- factor(result_sLDSC$Category, levels = annot_list)
```


```{r data table}
DT::datatable(format(result_sLDSC[,1:8], digits = 2), options = list(scrollX = TRUE, keys = TRUE, pageLength = length(annot_list)),rownames = F)
```


```{r plots}
ggplot_enrichment(result_sLDSC, title = "")
ggplot_log2_enrichment(result_sLDSC, title = "")
```

